---
title: "Summary of Activity"
author: "DZastera"
date: "August 7, 2016"
output: 
  html_document: 
    keep_md: yes
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load}
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
act <- read.csv("activity.csv")
```

## Total Steps Per Day

The total number of steps taken per day is summarized in this table:

```{r tot_steps}
tot_steps <- act %>% group_by(date) %>% summarise(Total = sum(steps, na.rm=TRUE))
print(data.frame(tot_steps), row.names=FALSE)
```

The historgram of steps per day is presented in this plot:
```{r tot_hist}
ggplot(tot_steps, aes(x=Total))+
  geom_histogram(bins=61)+
  labs(x = "Total Steps Per Day", y = "Count")
```

The mean number of steps per day is `r mean(tot_steps$Total)` and the median number of steps per day is `r median(tot_steps$Total)`.

## Average Daily Pattern

```{r step_ts}
step_ts <- act %>% group_by(interval) %>% summarise(avg=mean(steps, na.rm=TRUE))
ggplot(step_ts, aes(x=interval, y=avg))+
  geom_line()+
  labs(x = "Time Interval", y = "Mean Steps Per Interval")
```

The maximum average number of steps taken occurs in the `r which(step_ts==max(step_ts$avg))` interval.

## Analysis of NAs

The total number of missing values in the dataset is `r sum(is.na(act))`.

NAs in the dataset can be replaced by the interval mean with:

```{r impute}
imputed <- act %>% group_by(interval) %>% mutate(imp_step = mean(steps, na.rm=TRUE))
imputed$steps[is.na(imputed$steps)] <- imputed$imp_step[is.na(imputed$steps)]
```

Now, the total steps taken each day is:

```{r tot_steps_imp}
tot_steps_imp <- imputed %>% group_by(date) %>% summarise(Total = sum(steps, na.rm=TRUE))
print(data.frame(tot_steps_imp), row.names=FALSE)
```

The historgram of steps per day with the imputed data is:
```{r tot_hist_imp}
ggplot(tot_steps_imp, aes(x=Total))+
  geom_histogram(bins=61)+
  labs(x = "Total Steps Per Day", y = "Count")
```

The mean number of steps per day is `r mean(tot_steps_imp$Total)` and the median number of steps per day is `r median(tot_steps_imp$Total)`.

The imputed dataset has a larger number of mean steps per day and a larger median number of steps per day.


## Impact of Weekdays

```{r weekday}
act$weekday <- ifelse(weekdays(as.Date(act$date)) %in% c("Saturday","Sunday"),"weekend", "weekday") 
```

``` {r weekday_plot}
step_ts_weekday <- act %>% group_by(interval, weekday) %>% summarise(avg=mean(steps, na.rm=TRUE))
ggplot(step_ts_weekday, aes(x=interval, y=avg, colour=weekday))+
  geom_line()+
  facet_wrap(~weekday, ncol=1)+
  labs(x = "Time Interval", y = "Mean Steps Per Interval")+
  theme(legend.position="bottom")
```

